---
title: "ML Final Project"
author: "Sandra"
date: "12/18/2021"
output:
  pdf_document: default
  html_document: default
---


```{r}
rm(list=ls())
```


```{r}
library(psych)
library(MASS)
library(pROC)
library(rpart)
library(randomForest)
library(nnet)
library(e1071)
library(gam)
```

```{r}
# Data Preparation
labels = read.csv("training_set_labels.csv", header = TRUE)
df = read.csv("training_set_features.csv", header = TRUE)
data = merge(df, labels, by = "respondent_id")
data = data[,-c(12,20:22,28:29,31:32,35:36,38)]
head(data,5)

```

```{r}
describe(data)
# describe(test)
```
```{r}
# Split data into train and test dataset
set.seed(100)
trn <- runif(nrow(data)) < 0.7
train <- data[trn == TRUE,]
test <- data[trn == FALSE,]
train <- train[complete.cases(train), -c(1)]
test <- test[complete.cases(test), -c(1)]
```

```{r}
# Confusion Matrix 
TPR <- function(y,yhat)  { sum(y==1 & yhat==1) / sum(y==1) }
TNR <- function(y,yhat)  { sum(y==0 & yhat==0) / sum(y==0) }
ME <- function(y, yhat)  { (sum(y==1 & yhat==0)+sum(y==0 & yhat==1))/sum(y==1|y==0)}
```


*********************************
** PART I -- LINEAR REGRESSION **
*********************************
R-square:
The larger R-squared, the better the regression model fits your data. 
The R^2 for improved model m1 and original ols_full are similar to 0.33.
The low R^2 presents the model cannot fit the sample data well.

AIC:
The lower the value for AIC, the better the fit of the model. The absolute value of the AIC value is not important. It can be positive or negative.
When applying OLS regression, AIC is around -26848, which means it is quite nice to fit the model by using "backward selection" method.


```{r}
# OLS regression
ols_full <- lm(h1n1_vaccine ~ ., data = train)
summary(ols_full)
summary(ols_full)$adj.r.squared

# Enhance the model
m1 <- lm(h1n1_vaccine ~ h1n1_knowledge+doctor_recc_h1n1+child_under_6_months+
           health_worker+health_insurance+opinion_h1n1_vacc_effective+
           opinion_h1n1_risk+age_group+race+sex, data = train)
summary(m1)
summary(m1)$adj.r.squared

# Predictions
yhat.ols <- predict(m1, newdata = test)
table(test$h1n1_vaccine, (yhat.ols > 0.5))
TPR(test$h1n1_vaccine, (yhat.ols > 0.5))
TNR(test$h1n1_vaccine, (yhat.ols > 0.5))
ME(test$h1n1_vaccine, (yhat.ols > 0.5))
```


```{r}
# Rely on "Backward Selection" to obtain a model
ols_real_full <- lm(h1n1_vaccine ~., data = train)
step(ols_real_full, direction="backward")
```

*********************************
** PART II -- LOGISTIC REGRESSION **
*********************************
Logistic regression is used for modeling categorical outcomes, particularly no/yes, 0/1 outcomes.Such problems are called (binary) classification problems.
Under the consideration of dependent variable “h1n1_vaccine”, we decide to use logistic regression with binary classification. 

```{r}
# Apply glm() -- logistic regression model
glm <- glm(h1n1_vaccine ~.,family = "binomial", data = train)
summary(glm)
yhat_glm <- predict(glm, type="response")

# confusion matrix
table(train$h1n1_vaccine, (yhat_glm>0.5))
TPR(train$h1n1_vaccine,(yhat_glm>0.5))
TNR(train$h1n1_vaccine,(yhat_glm>0.5))
ME(train$h1n1_vaccine,(yhat_glm>0.5))

yhat.glm <- predict(glm, test, type="response")
table(test$h1n1_vaccine, (yhat.glm > 0.5))
TPR(test$h1n1_vaccine, (yhat.glm > 0.5))
TNR(test$h1n1_vaccine, (yhat.glm > 0.5))

# Misclassification Error
ME(test$h1n1_vaccine, (yhat.glm > 0.5))
```


```{r}
# LDA/QDA
lda <- lda(h1n1_vaccine ~ ., data=train)
yhat_lda <- predict(lda)$posterior[,2]
table(train$h1n1_vaccine, (yhat_lda >0.5))
yhat.lda <- predict(lda, test)$posterior[,2]
table(test$h1n1_vaccine, (yhat.lda > 0.5))
TPR(test$h1n1_vaccine, (yhat.lda >0.5))
TNR(test$h1n1_vaccine, (yhat.lda >0.5))
ME(test$h1n1_vaccine, (yhat.lda >0.5))

qda <- qda(h1n1_vaccine ~ ., data=train)
yhat_qda <- predict(qda)$posterior[,2]
table(train$h1n1_vaccine, (yhat_qda >0.5))
yhat.qda <- predict(qda, test)$posterior[,2]
table(test$h1n1_vaccine, (yhat.qda > 0.5))
TPR(test$h1n1_vaccine, (yhat.qda >0.5))
TNR(test$h1n1_vaccine, (yhat.qda >0.5))
ME(test$h1n1_vaccine, (yhat.qda >0.5))
```


```{r}
# Plot ROC graph
par(mfrow=c(1,1))
glm.roc <- roc(test$h1n1_vaccine, yhat.glm, direction = "<")
glm.roc
lda.roc <- roc(test$h1n1_vaccine, yhat.lda, direction = "<")
lda.roc
qda.roc <- roc(test$h1n1_vaccine, yhat.qda, direction = "<")
qda.roc
plot(glm.roc, lwd=3)
lines(lda.roc, lwd=3, col = "yellow")
lines(qda.roc, lwd=3, col = "blue")
legend("bottomright",title="ROC Curves",c("glm","lda","qda"), fill=c("black","yellow","blue"))
```

*********************************
** PART III -- Classification Tree **
*********************************
```{r}
paste(colnames(train[,-c(1)]), collapse = "+")
form1 <- formula(h1n1_vaccine ~ h1n1_concern+h1n1_knowledge+behavioral_antiviral_meds+behavioral_avoidance+behavioral_face_mask+behavioral_wash_hands+behavioral_large_gatherings+behavioral_outside_home+behavioral_touch_face+doctor_recc_h1n1+chronic_med_condition+child_under_6_months+health_worker+health_insurance+opinion_h1n1_vacc_effective+opinion_h1n1_risk+opinion_h1n1_sick_from_vacc+age_group+education+race+sex+income_poverty+employment_status+household_adults+household_children)

t1 <- rpart(form1, data=train, cp=.001, method="class")
plot(t1,uniform=T,compress=T,margin=.05,branch=0.9)
text(t1, cex=.4, col="navy",use.n=TRUE)

plotcp(t1) #plot cross-validation results
CP <- printcp(t1) #display cp table

cp <- CP[,1][CP[,4] == min(CP[,4])]
cp

t2 <- prune(t1,cp=cp[1])
plot(t2,uniform=T,compress=T,margin=.05,branch=0.3)
text(t2, cex=.7, col="navy",use.n=TRUE)

# Predictions
yhat.t2 <- predict(t2, test, type="prob")[,2]
table(test$h1n1_vaccine, (yhat.t2>0.5))
TPR(test$h1n1_vaccine,(yhat.t2>0.5))
TNR(test$h1n1_vaccine,(yhat.t2>0.5))
ME(test$h1n1_vaccine,(yhat.t2>0.5))
tree.roc <- roc(test$h1n1_vaccine, yhat.t2 , direction="<")
tree.roc
```


*********************************
** PART IV -- Random Forest **
*********************************
```{r}
train_rf <- train[,-c(18:23)]
X <- as.matrix(train_rf[,-c(20)])
Y <- factor(train_rf$h1n1_vaccine)
rf1 <- randomForest(x=X, y=Y, data=train_rf, ntree=500, mtry=3, importance=T, na.action=na.omit)
summary(rf1)

head(rf1$importance)

# variable importance ranking 1
varImpPlot(rf1, sort = TRUE , main = "Random Forest Importance Plot")

# variable importance ranking 2
imp <- rf1$importance[,4]
ord <- order(imp, decreasing=T)
imp <- imp[ord]

par(mar=c(2, 8, 4, 2) + 0.1)
barplot(imp, col='lavender', horiz=TRUE, las=1, cex.names=.8)
title("Random Forest Variable Importance Plot")

# Predictions
yhat.rf <- predict(rf1, test, type="prob")[,2]
table(test$h1n1_vaccine, (yhat.rf>0.5))
TPR(test$h1n1_vaccine,(yhat.rf>0.5))
TNR(test$h1n1_vaccine,(yhat.rf>0.5))
ME(test$h1n1_vaccine,(yhat.rf>0.5))

# Plot ROC graph
par(mfrow=c(1,1))
rf.roc <- roc(test$h1n1_vaccine, yhat.rf, direction="<")
rf.roc
plot(rf.roc, lwd=3, col = "green")
lines(tree.roc, lwd=3, col = "red")
lines(glm.roc, lwd=3)
legend("bottomright",title="ROC Curves",c("rf","tree","glm"), fill=c("green","red","black"))
```

*********************************
** PART V -- Neural Nets **
*********************************
```{r}
data_nn <- data[,-c(1,19:24)]
data_nn[,-c(20)] <- scale(data_nn[,-c(20)])
data_nn$h1n1_vaccine <- factor(data_nn$h1n1_vaccine)
train2 <- data_nn[trn==TRUE,]
test2 <- data_nn[trn==FALSE,]

paste(colnames(data_nn), collapse = "+")
form2 <- h1n1_vaccine ~ h1n1_concern+h1n1_knowledge+behavioral_antiviral_meds+behavioral_avoidance+behavioral_face_mask+behavioral_wash_hands+behavioral_large_gatherings+behavioral_outside_home+behavioral_touch_face+doctor_recc_h1n1+chronic_med_condition+child_under_6_months+health_worker+health_insurance+opinion_h1n1_vacc_effective+opinion_h1n1_risk+opinion_h1n1_sick_from_vacc+household_adults+household_children

n1 <- nnet(form2, data = train2, size = 7, maxit = 500, decay=0.002)

# Predictions
yhat.n1 <- predict(n1, test)
table(test$h1n1_vaccine, (yhat.n1>0.5))
TPR(test$h1n1_vaccine,(yhat.n1>0.5))
TNR(test$h1n1_vaccine,(yhat.n1>0.5))
ME(test$h1n1_vaccine,(yhat.n1>0.5))

# Plot ROC graph
par(mfrow=c(1,1))
nn.roc <- roc(test$h1n1_vaccine, yhat.n1, direction="<")
nn.roc
plot(nn.roc, lwd=3, col = "purple")
lines(rf.roc, lwd=3, col = "green")
lines(glm.roc, lwd=3)
legend("bottomright",title="ROC Curves",c("nnet","rf","glm"), fill=c("purple","green","black"))
```
















